Open Data, Open Source and Open Standards in chemistry - The Blue Obelisk five years on

### 2011-04-29 paper.tex

@@

An example of such an interaction was the donation by eMolecules, Inc.
to Open Babel of code for the canonicalisation of molecules and
fragments (Nov 2006). eMolecules is an online vendor of chemicals that
uses Open Babel under-the-hood to manage its compound collections.
Another more recent example occurred in July 2010 when Silicos, a
Belgian company that provides services
in the area of cheminformatics, released several command line
applications based on Open Babel as well as donating code to the
project. For example, the Pharao
tool released by them is a comprehensive solution for pharmacophore
searching that provides extensive support for a variety of
pharmacophore searches. Other tools released by them include Sieve for
filtering by molecular property and XXX.

@@

In the area of cheminformatics toolkits, two of the existing toolkits
Open Babel and RDKit are planning to work together on a common
underlying framework called MolCore. This project is still in the
planning stage, but if it is a success it will mean that the two
libraries will be interoperable (while retaining their existing focus)
but also that the cost of maintaining the code will be shared among
more developers, freeing time for the development of new features.

>>>>

Note that while some may advocate also for Open Access to
publications, the goals focus more on the availability of code (to
reproduce results), standards (to exchange data), and the scientific
data itself. All three of these goals stem from the basics of the
scientific method for data sharing and reproducibilty.

@@

An example of such an interaction was the donation by eMolecules, Inc.
to Open Babel of code for the canonicalisation of molecules and
fragments (Nov 2006). eMolecules is an online vendor of chemicals that
uses Open Babel under-the-hood to manage its compound collections.

Another more recent example occurred in July 2010 when Silicos, a
Belgian company that provides services
in the area of cheminformatics, released several command line
applications based on Open Babel as well as donating code to the
project. For example, the Pharao
tool released by them is a comprehensive solution for pharmacophore
searching that provides extensive support for a variety of
pharmacophore searches. Other tools released by Silicos include Sieve for
filtering by molecular property, Stripper for removing core scaffold
structures from a molecule set, and Piramid for molecular alignment
using shape determined by the Gaussian volumes as a descriptor.

@@

In the area of cheminformatics toolkits, two of the existing toolkits
Open Babel and RDKit are planning to work together on a common
underlying framework called MolCore.\cite{WebMolCore} This project is still in the
planning stage, but if it is a success it will mean that the the two
libraries will be interoperable (while retaining their existing focus)
but also that the cost of maintaining the code will be shared among
more developers, freeing time for the development of new features.
